Supplement - Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-09-25 submission/supplement.Rmd
description of supplementary tables

\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/Supplement_KDM5C_ChIPseq.pdf}
  \caption[KDM5C binds to a subset of germline RNA-seq differentially expressed genes]{\textbf{KDM5C binds to a subset of germline RNA-seq differentially expressed genes.} \textbf{A.} Bar graph of the number of germline-enriched DEGs with promoter KDM5C ChIP-seq peaks in wild-type EpiLCs (Top) and PNCs (Bottom). RNA-seq DEGs were classified as shared betwen EpiLCs and the brain (Common), unique to EpiLCs (EpiLC Only), or unique to one or multiple brain regions (Brain Only). \textbf{B.} Average bigwigs of two example RNA-seq DEGs dysregulated in EpiLCs but not the brain, \textit{Dazl} and \textit{Stra8}. Top is the RNA-seq tracks for wild-type (WT) and \textit{Kdm5c}-KO (5CKO) EpiLCs, bottom is the KDM5C ChIP-seq tracks, with the annotated transcription start site (TSS) for each gene. \textbf{C.} Same as B but for two example DEGs only dysregualted in the brain and not expressed in EpiLCs, \textit{Ccnb1ip1} and \textit{Spata18}. \textbf{D.} HOMER motif analysis of all KDM5C-unbound germline DEGs shows signficant enrichment of multiple RFX members and their X-box motif. \textbf{E.} KDM5C ChIP-seq shows no KDM5C accumulation at the \textit{Rfx2} promoter in EpiLCs or PNCs.}
  \label{figurelabel}
\end{figure}

\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/Supplement_WGBS.pdf}
  \caption[Loss of KDM5C impairs CpG methylation of germline gene promoters]{\textbf{[Loss of KDM5C impairs CpG methylation of germline gene promoters.} \textbf{A.} Volcano plot of whole genome bisulfite sequencing (WGBS) for all gene promoters in wild-type (WT) versus \textit{Kdm5c}-KO extended EpiLCs (exEpiLCs). Significantly differentially methylated promoters (q < 0.01, |methylation difference| > 25\%). Germline promoters highlighted in green, non-germline promoters in light blue, non-significant promoters in gray. \textbf{B.} enrichPlot gene ontology of all promoters significantly hypomethylated in \textit{Kdm5c}-KO exEpiLCs.}
  \label{figurelabel}
\end{figure}

>>>>

# Supplementary Tables

**Supplementary table 1: Misexpression of tissue-enriched genes within the _Kdm5c_-KO brain.** 1) DESeq2 results table of the _Kdm5c_-KO hippocampus, with tissue-enriched genes annotated 2) Same as 1 but for the _Kdm5c_-KO amygdala 3) Results of the Fisher exact test for enrichment tissue-specific genes. 4) enrichPlot gene ontology results of testis-enriched genes misexpressed in the amygdala or hippocampus.

**Supplementary table 2: Mouse germline-enriched genes.** List of mouse germline-enriched genes identified in this study (see Methods). Includes whether germline gene promoters have KDM5C ChIP-seq peaks and CpG islands.

**Supplementary table 3: Germline gene misexpression in _Kdm5c_ mutant EpiLCs.** 1) DESeq2 results table for male _Kdm5c_-KO (_Kdm5c^-/y^_) epiblast-like cells (EpiLCs), with annotations for tissue-enriched genes. 2) Results of Fisher exact test on XY EpiLC tissue-enriched genes. 3) enrichPlot gene ontology results of germline-enriched genes misexpressed in male _Kdm5c_-KO EpiLCs, amygdala, and hippocampus. 4) Germline genes misexpressed in XY _Kdm5c_-KO (_Kdm5c^-/y^_), XX _Kdm5c_-KO (_Kdm5c^-/-^_), and XX _Kdm5c_-HET (_Kdm5c^-/+^_) EpiLCs. 5) enrichPlot gene ontology results of germline genes misexpressed in male versus female _Kdm5c_ mutant EpiLCs

**Supplementary table 4: KDM5C ChIP-seq in EpiLCs and PNCs.** 1) enrichPlot gene ontology results of KDM5C-bound promoters (see Methods) in epiblast-like cells (EpiLCs) and forebrain primary neuron cultures (PNCs). 2) HOMER known motif analysis of germline gene promoters (TSS &#177; 500 bp) bound by KDM5C . 3) HOMER known motif analysis of germline genes not bound by KDM5C.

**Supplementary table 5: Germline gene CpG islands and promoter CpG methylation.** 1) enrichPlot gene ontology results of germline genes with and without CpG islands (CGIs) within their promoter (TSS &#177; 500 bp). 2) methylKit whole genome bisulfite sequencing (WGBS) results table comparing germline gene promoter CpG methylation (TSS &#177; 500 bp) in male wild-type ESCs versus wild-type extended EpiLCs (exEpiLCs). 3) methylKit WGBS results table comparing germline gene promoter CpG methylation in male _Kdm5c_-KO versus wild-type exEpiLCs.

# Figures
\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/Supplement_KDM5C_ChIPseq.pdf}
  \caption[KDM5C binds to a subset of germline RNA-seq differentially expressed genes]{\textbf{KDM5C binds to a subset of germline RNA-seq differentially expressed genes.} \textbf{A.} Bar graph of the number of germline-enriched DEGs with promoter KDM5C ChIP-seq peaks in wild-type EpiLCs (Top) and PNCs (Bottom). RNA-seq DEGs were classified as shared betwen EpiLCs and the brain (Common), unique to EpiLCs (EpiLC Only), or unique to one or multiple brain regions (Brain Only). \textbf{B.} Average bigwigs of two example RNA-seq DEGs dysregulated in EpiLCs but not the brain, \textit{Dazl} and \textit{Stra8}. Top is the RNA-seq tracks for wild-type (WT) and \textit{Kdm5c}-KO (5CKO) EpiLCs, bottom is the KDM5C ChIP-seq tracks, with the annotated transcription start site (TSS) for each gene. \textbf{C.} Same as B but for two example DEGs only dysregualted in the brain and not expressed in EpiLCs, \textit{Ccnb1ip1} and \textit{Spata18}. \textbf{D.} HOMER motif analysis of all KDM5C-unbound germline DEGs shows signficant enrichment of multiple RFX members and their X-box motif. \textbf{E.} KDM5C ChIP-seq shows no KDM5C accumulation at the \textit{Rfx2} promoter in EpiLCs or PNCs.}
  \label{figurelabel}
\end{figure}

\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/Supplement_WGBS.pdf}
  \caption[Loss of KDM5C impairs CpG methylation of germline gene promoters]{\textbf{Loss of KDM5C impairs CpG methylation of germline gene promoters.} \textbf{A.} Volcano plot of whole genome bisulfite sequencing (WGBS) for all gene promoters in wild-type (WT) versus \textit{Kdm5c}-KO extended EpiLCs (exEpiLCs). Significantly differentially methylated promoters (q < 0.01, |methylation difference| > 25\%). Germline promoters highlighted in green, non-germline promoters in light blue, non-significant promoters in gray. \textbf{B.} enrichPlot gene ontology of all promoters significantly hypomethylated in \textit{Kdm5c}-KO exEpiLCs.}
  \label{figurelabel}
\end{figure}
